---
title: "02-groupfacAnalysis_BetweenWithinNetwork"
output: html_document
---


```{r}
rm(list=ls())
setwd("~/Documents/oleary/rest/")
```


```{r}
library(dplyr)
library(car)
library(lsr)
library(ggplot2)
library(ggstatsplot)
library(foreign)
library(psych)
library(Hmisc)
library(ppcor)
```


```{r}
finaldf <- read.table("finaldf.tsv", header = T)
finaldf$groupfac <- factor(finaldf$groupfac,levels(finaldf$groupfac)[c(1,5,2,4,3)])
```


```{r}
p_anova <- rep (0, 91)

for (i in 4:94) {
    model <- lm(unlist(finaldf[i]) ~ groupfac, data=finaldf)
    anova <- Anova(model, contrasts = c("contr.sum", "contr.poly"), type="III")
    p_anova[i-3] <- anova$"Pr(>F)"[2]
    cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
    print(anova)
    print(etaSquared(model, type = 3, anova = FALSE))
    print(leveneTest(model))
    print(shapiro.test(model[["residuals"]]))
    hist(residuals(model), col="darkgray")
    plot(fitted(model),residuals(model))
}

```


```{r}
p_fdr <- p.adjust(p_anova, method = c("fdr"), n = length(p_anova))
p_fdr
```


```{r}
which(p_anova < 0.05)
p_anova[which(p_anova < 0.05)]
```

```{r}
names(finaldf)[8+3]
names(finaldf)[69+3]
```


```{r}
ggbetweenstats(data = finaldf, x = groupfac, y = Reward, plot.type = "boxviolin", type = "parametric",
               pairwise.comparisons = T, pairwise.annotation = "p.value", pairwise.display = "significant",
               p.adjust.method = "fdr", partial = T, effsize.type = "d", conf.level = 0.95, package = "ggsci", 
               palette = "nrc_npg", k = 2, messages = FALSE, bf.message = T, 
               title = "Within Network Connectivity - Reward", ylab = "Correlation") 

```


```{r}
ggbetweenstats(data = finaldf, x = groupfac, y = MedialTemporal.VentralAttention, plot.type = "boxviolin", 
               type = "parametric", pairwise.comparisons = T, pairwise.annotation = "p.value", 
               pairwise.display = "significant", p.adjust.method = "fdr", partial = T, 
               effsize.type = "d", conf.level = 0.95, package = "ggsci", 
               palette = "nrc_npg", k = 2, messages = FALSE, bf.message = T, 
               title = "Bwt Network: MedialTemporal - VentralAttention", ylab = "Correlation") 
```


```{r}
spss <- read.spss("../Complete_Dataset_fix_missing_values.sav", to.data.frame = T, stringsAsFactors = F);
```


```{r}
finaldf$DrinkingCategory <- spss$Drinking_Category
finaldf$BingeHistory <- as.factor(ifelse(finaldf$DrinkingCategory == "Binge                ", "Binge", 
                                  ifelse(finaldf$DrinkingCategory == "Binge History        ", "BingeHistory",
                                        NA)))

finaldf$ExtremeBingeHistory <- as.factor(ifelse(finaldf$DrinkingCategory == "Extreme Binge        ",
                                                "ExtremeBinge", 
                                         ifelse(finaldf$DrinkingCategory == "Extreme Binge History",
                                                "ExtremeBingeHistory", NA)))
```


# Binge History

```{r}
p_ttest_binge <- rep(0, 91)

for (i in 4:94) {
  bartlett_binge <- bartlett.test(unlist(finaldf[i]) ~ BingeHistory, data=finaldf)
  if (bartlett_binge$p.value >= 0.05) {
    ttest_binge <- t.test(unlist(finaldf[i]) ~ BingeHistory, data=finaldf, var.equal=TRUE, conf.level=0.95)} 
  else {ttest_binge <- t.test(unlist(finaldf[i]) ~ BingeHistory, data=finaldf, var.equal=FALSE, conf.level=0.95)}
  p_ttest_binge[i-3] <- ttest_binge$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_binge)
}
```


```{r}
p_fdr_ttestbinge <- p.adjust(p_ttest_binge, method = c("fdr"), n = length(p_ttest_binge))
p_fdr_ttestbinge
```



```{r}
p_ttest_binge[p_ttest_binge < 0.05]
which(p_ttest_binge < 0.05)
```

```{r}
names(finaldf)[1+3]
names(finaldf)[4+3]
names(finaldf)[16+3]
```

# Extreme Binge History

```{r}
p_ttest_exbinge <- rep (0, 91)

for (i in 4:94) {
  bartlett_exbinge <- bartlett.test(unlist(finaldf[i]) ~ ExtremeBingeHistory, data=finaldf)
  if (bartlett_exbinge$p.value >= 0.05) {
    ttest_exbinge <- t.test(unlist(finaldf[i]) ~ ExtremeBingeHistory, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_exbinge <- t.test(unlist(finaldf[i]) ~ ExtremeBingeHistory, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_ttest_exbinge[i-3] <- ttest_exbinge$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_exbinge)
}
```


```{r}
p_fdr_ttestexbinge <- p.adjust(p_ttest_exbinge, method = c("fdr"), n = length(p_ttest_exbinge))
p_fdr_ttestexbinge
```


```{r}
p_ttest_exbinge[p_ttest_exbinge < 0.05]
which(p_ttest_exbinge < 0.05)
```

```{r}
names(finaldf)[2+3]
names(finaldf)[56+3]
```

# YAACQ

```{r}
finaldf$V1_YAACQ_Social_Interpersonal <- spss$V1_YAACQ_Social_Interpersonal
finaldf$V1_YAACQ_Impaired_Control <- spss$V1_YAACQ_Impaired_Control
finaldf$V1_YAACQ_Self_Perception <- spss$V1_YAACQ_Self_Perception
finaldf$V1_YAACQ_Self_Care <- spss$V1_YAACQ_Self_Care
finaldf$V1_YAACQ_Risk_Taking <- spss$V1_YAACQ_Risk_Taking
finaldf$V1_YAACQ_Academic_Occupational <- spss$V1_YAACQ_Academic_Occupational
finaldf$V1_YAACQ_Dependence <- spss$V1_YAACQ_Dependence
finaldf$V1_YAACQ_Blackouts <- spss$V1_YAACQ_Blackouts
finaldf$V1_YAACQ_Total <- finaldf$V1_YAACQ_Social_Interpersonal +
  finaldf$V1_YAACQ_Impaired_Control +
  finaldf$V1_YAACQ_Self_Perception +
  finaldf$V1_YAACQ_Self_Care +
  finaldf$V1_YAACQ_Risk_Taking +
  finaldf$V1_YAACQ_Academic_Occupational +
  finaldf$V1_YAACQ_Dependence +
  finaldf$V1_YAACQ_Blackout

for (i in 98:106) {
  hist(finaldf[i])
}
  
```


```{r}
cor.fconYAACQ <- rcorr(as.matrix(finaldf[c(4:94, 98:106)]), type = c("spearman"))
write.table(cor.fconYAACQ$P, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fconYAACQcorrel-p-uncorrect.tsv")
write.table(cor.fconYAACQ$r, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fconYAACQcorrel-coeffi.tsv")

to.upper<-function(X) t(X)[lower.tri(X,diag=F)]
p.uncorrect.vector <- to.upper(cor.fconYAACQ$P)

p.fdr.vector <- p.adjust(p.uncorrect.vector, method = c("fdr"), n = length(p.uncorrect.vector))

p.fdr.matrix <- matrix(0, 100, 100)
p.fdr.matrix[lower.tri(p.fdr.matrix, diag=FALSE)] <- p.fdr.vector
p.fdr.matrix <- as.data.frame(t(p.fdr.matrix))
rownames(p.fdr.matrix) <- rownames(cor.fconYAACQ$P)
colnames(p.fdr.matrix) <- colnames(cor.fconYAACQ$P)

write.table(p.fdr.matrix, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "~/Documents/oleary/rest/fconYAACQcorrel-p-correct.tsv")
```


# Binge/ExBinge Frequency

```{r}
myvars <- c(
"Subject_ID",
"DaysBingeDrinking_Female_6thTo8thGrade",
"DaysBingeDrinking_Female_9thGrade",
"DaysBingeDrinking_Female_10thGrade",
"DaysBingeDrinking_Female_11thGrade",
"DaysBingeDrinking_Female_12thGrade",
"DaysBingeDrinking_Female_SummerAfterHighSchool",
"DaysBingeDrinking_Female_FreshmanYear",
"DaysBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysBingeDrinking_Female_ThisAcademicYear",
"V1_DaysBingeDrinking_Female_Past30Days",

"DaysBingeDrinking_Male_6thTo8thGrade",
"DaysBingeDrinking_Male_9thGrade",
"DaysBingeDrinking_Male_10thGrade",
"DaysBingeDrinking_Male_11thGrade",
"DaysBingeDrinking_Male_12thGrade",
"DaysBingeDrinking_Male_SummerAfterHighSchool",
"DaysBingeDrinking_Male_FreshmanYear",
"DaysBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysBingeDrinking_Male_ThisAcademicYear",
"V1_DaysBingeDrinking_Male_Past30Days",

"DaysExtremeBingeDrinking_Female_6thTo8thGrade",
"DaysExtremeBingeDrinking_Female_9thGrade",
"DaysExtremeBingeDrinking_Female_10thGrade",
"DaysExtremeBingeDrinking_Female_11thGrade",
"DaysExtremeBingeDrinking_Female_12thGrade",
"DaysExtremeBingeDrinking_Female_SummerAfterHighSchool",
"DaysExtremeBingeDrinking_Female_FreshmanYear",
"DaysExtremeBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Female_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Female_Past30Days",

"DaysExtremeBingeDrinking_Male_6thTo8thGrade",
"DaysExtremeBingeDrinking_Male_9thGrade",
"DaysExtremeBingeDrinking_Male_10thGrade",
"DaysExtremeBingeDrinking_Male_11thGrade",
"DaysExtremeBingeDrinking_Male_12thGrade",
"DaysExtremeBingeDrinking_Male_SummerAfterHighSchool",
"DaysExtremeBingeDrinking_Male_FreshmanYear",
"DaysExtremeBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Male_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Male_Past30Days",

"TotalDaysUsedCannabis_6thTo8thGrade",
"TotalDaysUsedCannabis_9thGrade",
"TotalDaysUsedCannabis_10thGrade",
"TotalDaysUsedCannabis_11thGrade",
"TotalDaysUsedCannabis_12thGrade",
"TotalDaysUsedCannabis_SummerAfterHighSchool",
"TotalDaysUsedCannabis_FreshmanYear",
"TotalDaysUsedCannabis_SummerAfterFreshmanYear",
"TotalDaysUsedCannabis_ThisAcademicYear",
"V1_TotalDaysUsedCannabis_Past30Days",

"AvgAmountCannabisSmoked_6thTo8thGrade",
"AvgAmountCannabisSmoked_9thGrade",
"AvgAmountCannabisSmoked_10thGrade",
"AvgAmountCannabisSmoked_11thGrade",
"AvgAmountCannabisSmoked_12thGrade",
"AvgAmountCannabisSmoked_SummerAfterHighSchool",
"AvgAmountCannabisSmoked_FreshmanYear",
"AvgAmountCannabisSmoked_SummerAfterFreshmanYear",
"AvgAmountCannabisSmoked_ThisAcademicYear",
"V1_AvgAmountCannabisSmoked_Past30Days",

"TotalDaysUsedTobacco_6thTo8thGrade",
"TotalDaysUsedTobacco_9thGrade",
"TotalDaysUsedTobacco_10thGrade",
"TotalDaysUsedTobacco_11thGrade",
"TotalDaysUsedTobacco_12thGrade",
"TotalDaysUsedTobacco_SummerAfterHighSchool",
"TotalDaysUsedTobacco_FreshmanYear",
"TotalDaysUsedTobacco_SummerAfterFreshmanYear",
"TotalDaysUsedTobacco_ThisAcademicYear",
"V1_TotalDaysUsedTobacco_Past30Days",

"AvgAmountTobaccoSmoked_6thTo8thGrade",
"AvgAmountTobaccoSmoked_9thGrade",
"AvgAmountTobaccoSmoked_10thGrade",
"AvgAmountTobaccoSmoked_11thGrade",
"AvgAmountTobaccoSmoked_12thGrade",
"AvgAmountTobaccoSmoked_SummerAfterHighSchool",
"AvgAmountTobaccoSmoked_FreshmanYear",
"AvgAmountTobaccoSmoked_SummerAfterFreshmanYear",
"AvgAmountTobaccoSmoked_ThisAcademicYear",
"V1_AvgAmountTobaccoSmoked_Past30Days")

substance <- spss[myvars]


  
```

```{r}
cor.fcondrug <- rcorr(as.matrix(finaldf[c(4:94, 115:118)]), type = c("spearman"))
write.table(cor.fcondrug$P, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fcondrugcorrel-p-uncorrect.tsv")
write.table(cor.fcondrug$r, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fcondrugcorrel-coeffi.tsv")

to.upper<-function(X) t(X)[lower.tri(X,diag=F)]
p.uncorrect.vector.drug <- to.upper(cor.fcondrug$P)

p.fdr.vector.drug <- p.adjust(p.uncorrect.vector.drug, method = c("fdr"), n = length(p.uncorrect.vector.drug))

p.fdr.matrix.drug <- matrix(0, 95, 95)
p.fdr.matrix.drug[lower.tri(p.fdr.matrix.drug, diag=FALSE)] <- p.fdr.vector.drug
p.fdr.matrix.drug <- as.data.frame(t(p.fdr.matrix.drug))
rownames(p.fdr.matrix.drug) <- rownames(cor.fcondrug$P)
colnames(p.fdr.matrix.drug) <- colnames(cor.fcondrug$P)

write.table(p.fdr.matrix.drug, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fcondrugcorrel-p-correct.tsv")
```

# FCON and YAACQ correl without controls

```{r}
finaldf$V1_Age <- as.numeric(as.character(spss$V1_Age))
finaldf$Sub_Gender.1 <- spss$Sub_Gender.1
# 2 == Female
finaldf$Gender <- as.numeric(finaldf$Sub_Gender.1)
```


```{r}
finaldf_excon <- finaldf %>% filter(groupfac != "Control")

for (i in 98:106) {
  hist(finaldf_excon[i])}

```

```{r}
cor.fconYAACQ.fullexcon <- rcorr(as.matrix(finaldf[c(4:94, 98:106)]), type = c("spearman"))
write.table(cor.fconYAACQ.fullexcon$P, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "FULLfconYAACQcorrel-p-uncorrectEXCON.tsv")
write.table(cor.fconYAACQ.fullexcon$r, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "FULLfconYAACQcorrel-coeffiEXCON.tsv")

to.upper<-function(X) t(X)[lower.tri(X,diag=F)]
p.uncorrect.vector <- to.upper(cor.fconYAACQ.fullexcon$P)

p.fdr.vector <- p.adjust(p.uncorrect.vector, method = c("fdr"), n = length(p.uncorrect.vector))

p.fdr.matrix <- matrix(0, 100, 100)
p.fdr.matrix[lower.tri(p.fdr.matrix, diag=FALSE)] <- p.fdr.vector
p.fdr.matrix <- as.data.frame(t(p.fdr.matrix))
rownames(p.fdr.matrix) <- rownames(cor.fconYAACQ.fullexcon$P)
colnames(p.fdr.matrix) <- colnames(cor.fconYAACQ.fullexcon$P)

write.table(p.fdr.matrix, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "~/Documents/oleary/rest/FULLfconYAACQcorrel-p-correctEXCON.tsv")
```

` partial.r`
The second usage is to partial a set of variables(y) out of another set (x). It is sometimes convenient to partial the effect of a number of variables (e.g., sex, age, education) out of the correlations of another set of variables. This could be done laboriously by finding the residuals of various multiple correlations, and then correlating these residuals. The matrix algebra alternative is to do it directly. To find the confidence intervals and "significance" of the correlations, use the corr.p function with n = n - s where s is the numer of covariates.
IN THIS CASE N = 181 - 2 = 179

```{r}
finaldf_excon <- data.matrix(finaldf_excon)

cor.fconYAACQ.excon <- partial.r(finaldf_excon, c(4:94, 98:106), c(107,109), method="spearman")

cor.fconYAACQ.excon.p <-  corr.p(cor.fconYAACQ.excon, n = 179, adjust="fdr", alpha=.05, minlength=2)

write.table(cor.fconYAACQ.excon, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fconYAACQcorrel-coeffi-EXCON.tsv")
  
write.table(cor.fconYAACQ.excon.p$p, row.names = T, col.names = NA, sep = "\t", 
            quote = F, file = "fconYAACQcorrel-p-correct-EXCON.tsv")

```

# Control vs. Binge 

```{r}
finaldf$twogroup <- ifelse(finaldf$groupfac == "Control", "Control", "Binge") 
```


```{r}
p_ttest_2group <- rep (0, 91)

for (i in 4:94) {
  bartlett_2group <- bartlett.test(unlist(finaldf[i]) ~ twogroup, data=finaldf)
  if (bartlett_2group$p.value >= 0.05) {
    ttest_2group <- t.test(unlist(finaldf[i]) ~ twogroup, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_2group <- t.test(unlist(finaldf[i]) ~ twogroup, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_ttest_2group[i-3] <- ttest_2group$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_2group)
}
```


```{r}
p_fdr_ttest2group<- p.adjust(p_ttest_2group, method = c("fdr"), n = length(p_ttest_2group))
p_fdr_ttest2group
```


```{r}
p_ttest_2group[p_ttest_2group < 0.05]
which(p_ttest_2group < 0.05)
```




```{r}
write.table(finaldf, row.names = F, col.names = T, sep = "\t", quote = F, file = "finaldf_fin.tsv")
```


