---
title: "04- Sex and Age Effects"
output: html_document
---


```{r}
rm(list=ls())
setwd("~/Documents/oleary/rest/")
```


```{r}
library(dplyr)
library(car)
library(lsr)
library(ggplot2)
library(ggstatsplot)
library(foreign)
library(psych)
library(Hmisc)
library(ppcor)
```


```{r}
finaldf <- read.table("finaldf_fin.tsv", sep = "\t", header = T)
finaldf$groupfac <- factor(finaldf$groupfac,levels(finaldf$groupfac)[c(1,5,2,4,3)])
```


# Sex effects: Functional connectivity, YAACQ, drug use frequency


```{r}
p_sex <- rep (0, 112)

for (i in c(4:115)) {
  bartlett_sex <- bartlett.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf)
  if (bartlett_sex$p.value >= 0.05) {
    ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_sex[i-3] <- ttest_sex$p.value
  #cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  #print(ttest_sex)
}
```


```{r}
p_fdr_sex <- p.adjust(p_sex, method = c("fdr"), n = length(p_sex))
p_fdr_sex
```


```{r}
which(p_fdr_sex < 0.05)
p_sex[p_sex < 0.05]
which(p_sex < 0.05)
```

- Nothing survive FDR (although YAACQ subscores do not require correction -- I think?)

## Print out significant uncorrected p + significant variables' names

```{r}
names(finaldf)[which(p_sex < 0.05)+3]
```

## Print out the ttests that are interesting

```{r}
for (i in which(names(finaldf) == "V1_YAACQ_Self_Care")) {
  bartlett_sex <- bartlett.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf)
  if (bartlett_sex$p.value >= 0.05) {
    ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_sex[i-3] <- ttest_sex$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_sex)
}
```

Females showed higher negative consequences (self care)

```{r}
for (i in which(names(finaldf) == "DaysBinge_BeforeCollege")) {
  bartlett_sex <- bartlett.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf)
  if (bartlett_sex$p.value >= 0.05) {
    ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_sex[i-3] <- ttest_sex$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_sex)
}
```

Before college, females reported higher binge drink


```{r}
for (i in which(names(finaldf) == "DaysBinge_College")) {
  bartlett_sex <- bartlett.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf)
  if (bartlett_sex$p.value >= 0.05) {
    ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                            var.equal=TRUE, conf.level=0.95)} 
  else {ttest_sex <- t.test(unlist(finaldf[i]) ~ Sub_Gender.1, data=finaldf, 
                                var.equal=FALSE, conf.level=0.95)}
  p_sex[i-3] <- ttest_sex$p.value
  cat(paste('\nDependent variable:', names(finaldf[i]), '\n'))
  print(ttest_sex)
}
```


# Age effects: Functional connectivity, YAACQ, drug use frequency


```{r}
p_age <- rep (0, 112)

for (i in c(4:115)) {
    model <- lm(unlist(finaldf[i]) ~ V1_Age, data=finaldf)
    anova <- Anova(model, contrasts = c("contr.sum", "contr.poly"), type="III")
    p_age[i-3] <- anova$"Pr(>F)"[2]
    #cat(paste('\nDependent variable:', names(roi2roidf[i]), '\n'))
    #print(anova)
    #print(etaSquared(model, type = 3, anova = FALSE))
    #print(leveneTest(model))
    #print(shapiro.test(model[["residuals"]]))
    #hist(residuals(model), col="darkgray")
    #plot(fitted(model),residuals(model))
}
```


```{r}
p_fdr_age <- p.adjust(p_age, method = c("fdr"), n = length(p_age))
p_fdr_age
```

DaysBinge_College showed age effect after FDR correction

```{r}
names(finaldf[which(p_fdr_age < 0.05) + 3])
```

## Print out significant uncorrected p + significant variables' names

```{r}
p_age[p_age < 0.05]
which(p_age < 0.05)
```

```{r}
names(finaldf)[which(p_age < 0.05)+3]
```

