---
title: "CCA - 91 rsFC and drug use + YACCQ"
output:
  html_notebook:
    df_print: paged
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3
---

# The code is modified from Dinga et al
https://github.com/dinga92/niclin2019-biotypes

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=F)
rm(list=ls())
setwd("~/Documents/oleary/rest/")
```

# 0. Data loading and preparation

## Load data
### Resting state functional connectivity

```{r}
finaldf <- read.table("finaldf.tsv", header = T)
rsFC <- finaldf[c(1, 4:94)]

saveRDS(object = rsFC, file = "~/Documents/oleary/rest/91rsFC.rds")
```

### Covariates

```{r warning=FALSE}
library(foreign)
spss <- read.spss("../Complete_Dataset_fix_missing_values.sav", to.data.frame = T, stringsAsFactors = F)

finaldf$age <- as.numeric(as.character(spss$V1_Age))
finaldf$sex <- spss$Sub_Gender.1
covariates <- finaldf[c(1,95:96)]

saveRDS(object = covariates, file = "~/Documents/oleary/rest/covariates.rds")
```

### YAACQ

```{r}
finaldf$YAACQ_Social_Interpersonal <- spss$V1_YAACQ_Social_Interpersonal
finaldf$YAACQ_Impaired_Control <- spss$V1_YAACQ_Impaired_Control
finaldf$YAACQ_Self_Perception <- spss$V1_YAACQ_Self_Perception
finaldf$YAACQ_Self_Care <- spss$V1_YAACQ_Self_Care
finaldf$YAACQ_Risk_Taking <- spss$V1_YAACQ_Risk_Taking
finaldf$YAACQ_Academic_Occupational <- spss$V1_YAACQ_Academic_Occupational
finaldf$YAACQ_Dependence <- spss$V1_YAACQ_Dependence
finaldf$YAACQ_Blackouts <- spss$V1_YAACQ_Blackouts
finaldf$YAACQ_Total <- 
  finaldf$YAACQ_Social_Interpersonal +
  finaldf$YAACQ_Impaired_Control +
  finaldf$YAACQ_Self_Perception +
  finaldf$YAACQ_Self_Care +
  finaldf$YAACQ_Risk_Taking +
  finaldf$YAACQ_Academic_Occupational +
  finaldf$YAACQ_Dependence +
  finaldf$YAACQ_Blackout

yaacq <- finaldf[c(1,97:104)]
saveRDS(object = yaacq, file = "~/Documents/oleary/rest/yaacq.rds")
```

### Substance Use Variables

```{r}
myvars <- c(
"Subject_ID",
"DaysBingeDrinking_Female_6thTo8thGrade",
"DaysBingeDrinking_Female_9thGrade",
"DaysBingeDrinking_Female_10thGrade",
"DaysBingeDrinking_Female_11thGrade",
"DaysBingeDrinking_Female_12thGrade",
"DaysBingeDrinking_Female_SummerAfterHighSchool",
"DaysBingeDrinking_Female_FreshmanYear",
"DaysBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysBingeDrinking_Female_ThisAcademicYear",
"V1_DaysBingeDrinking_Female_Past30Days",

"DaysBingeDrinking_Male_6thTo8thGrade",
"DaysBingeDrinking_Male_9thGrade",
"DaysBingeDrinking_Male_10thGrade",
"DaysBingeDrinking_Male_11thGrade",
"DaysBingeDrinking_Male_12thGrade",
"DaysBingeDrinking_Male_SummerAfterHighSchool",
"DaysBingeDrinking_Male_FreshmanYear",
"DaysBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysBingeDrinking_Male_ThisAcademicYear",
"V1_DaysBingeDrinking_Male_Past30Days",

"DaysExtremeBingeDrinking_Female_6thTo8thGrade",
"DaysExtremeBingeDrinking_Female_9thGrade",
"DaysExtremeBingeDrinking_Female_10thGrade",
"DaysExtremeBingeDrinking_Female_11thGrade",
"DaysExtremeBingeDrinking_Female_12thGrade",
"DaysExtremeBingeDrinking_Female_SummerAfterHighSchool",
"DaysExtremeBingeDrinking_Female_FreshmanYear",
"DaysExtremeBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Female_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Female_Past30Days",

"DaysExtremeBingeDrinking_Male_6thTo8thGrade",
"DaysExtremeBingeDrinking_Male_9thGrade",
"DaysExtremeBingeDrinking_Male_10thGrade",
"DaysExtremeBingeDrinking_Male_11thGrade",
"DaysExtremeBingeDrinking_Male_12thGrade",
"DaysExtremeBingeDrinking_Male_SummerAfterHighSchool",
"DaysExtremeBingeDrinking_Male_FreshmanYear",
"DaysExtremeBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Male_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Male_Past30Days",

"TotalDaysUsedCannabis_6thTo8thGrade",
"TotalDaysUsedCannabis_9thGrade",
"TotalDaysUsedCannabis_10thGrade",
"TotalDaysUsedCannabis_11thGrade",
"TotalDaysUsedCannabis_12thGrade",
"TotalDaysUsedCannabis_SummerAfterHighSchool",
"TotalDaysUsedCannabis_FreshmanYear",
"TotalDaysUsedCannabis_SummerAfterFreshmanYear",
"TotalDaysUsedCannabis_ThisAcademicYear",
"V1_TotalDaysUsedCannabis_Past30Days",

"AvgAmountCannabisSmoked_6thTo8thGrade",
"AvgAmountCannabisSmoked_9thGrade",
"AvgAmountCannabisSmoked_10thGrade",
"AvgAmountCannabisSmoked_11thGrade",
"AvgAmountCannabisSmoked_12thGrade",
"AvgAmountCannabisSmoked_SummerAfterHighSchool",
"AvgAmountCannabisSmoked_FreshmanYear",
"AvgAmountCannabisSmoked_SummerAfterFreshmanYear",
"AvgAmountCannabisSmoked_ThisAcademicYear",
"V1_AvgAmountCannabisSmoked_Past30Days",

"TotalDaysUsedTobacco_6thTo8thGrade",
"TotalDaysUsedTobacco_9thGrade",
"TotalDaysUsedTobacco_10thGrade",
"TotalDaysUsedTobacco_11thGrade",
"TotalDaysUsedTobacco_12thGrade",
"TotalDaysUsedTobacco_SummerAfterHighSchool",
"TotalDaysUsedTobacco_FreshmanYear",
"TotalDaysUsedTobacco_SummerAfterFreshmanYear",
"TotalDaysUsedTobacco_ThisAcademicYear",
"V1_TotalDaysUsedTobacco_Past30Days",

"AvgAmountTobaccoSmoked_6thTo8thGrade",
"AvgAmountTobaccoSmoked_9thGrade",
"AvgAmountTobaccoSmoked_10thGrade",
"AvgAmountTobaccoSmoked_11thGrade",
"AvgAmountTobaccoSmoked_12thGrade",
"AvgAmountTobaccoSmoked_SummerAfterHighSchool",
"AvgAmountTobaccoSmoked_FreshmanYear",
"AvgAmountTobaccoSmoked_SummerAfterFreshmanYear",
"AvgAmountTobaccoSmoked_ThisAcademicYear",
"V1_AvgAmountTobaccoSmoked_Past30Days")

substance <- spss[myvars]
substance[substance[,] == -9] <- NA

substance$DaysBinge_BeforeCollege <- rowSums(substance[c(
"DaysBingeDrinking_Female_6thTo8thGrade",
"DaysBingeDrinking_Female_9thGrade",
"DaysBingeDrinking_Female_10thGrade",
"DaysBingeDrinking_Female_11thGrade",
"DaysBingeDrinking_Female_12thGrade",
"DaysBingeDrinking_Female_SummerAfterHighSchool",
"DaysBingeDrinking_Male_6thTo8thGrade",
"DaysBingeDrinking_Male_9thGrade",
"DaysBingeDrinking_Male_10thGrade",
"DaysBingeDrinking_Male_11thGrade",
"DaysBingeDrinking_Male_12thGrade",
"DaysBingeDrinking_Male_SummerAfterHighSchool"
)], na.rm = T)

substance$DaysExtremeBinge_BeforeCollege <- rowSums(substance[c(
"DaysExtremeBingeDrinking_Female_6thTo8thGrade",
"DaysExtremeBingeDrinking_Female_9thGrade",
"DaysExtremeBingeDrinking_Female_10thGrade",
"DaysExtremeBingeDrinking_Female_11thGrade",
"DaysExtremeBingeDrinking_Female_12thGrade",
"DaysExtremeBingeDrinking_Female_SummerAfterHighSchool",
"DaysExtremeBingeDrinking_Male_6thTo8thGrade",
"DaysExtremeBingeDrinking_Male_9thGrade",
"DaysExtremeBingeDrinking_Male_10thGrade",
"DaysExtremeBingeDrinking_Male_11thGrade",
"DaysExtremeBingeDrinking_Male_12thGrade",
"DaysExtremeBingeDrinking_Male_SummerAfterHighSchool"
)], na.rm = T)  

substance$DaysUsedCannabis_BeforeCollege <- rowSums(substance[c(
"TotalDaysUsedCannabis_6thTo8thGrade",
"TotalDaysUsedCannabis_9thGrade",
"TotalDaysUsedCannabis_10thGrade",
"TotalDaysUsedCannabis_11thGrade",
"TotalDaysUsedCannabis_12thGrade",
"TotalDaysUsedCannabis_SummerAfterHighSchool"
)], na.rm = T)

substance$DaysUsedCannabis_BeforeCollege <- rowSums(substance[c(
"TotalDaysUsedCannabis_6thTo8thGrade",
"TotalDaysUsedCannabis_9thGrade",
"TotalDaysUsedCannabis_10thGrade",
"TotalDaysUsedCannabis_11thGrade",
"TotalDaysUsedCannabis_12thGrade",
"TotalDaysUsedCannabis_SummerAfterHighSchool"
)], na.rm = T)

substance$DaysUsedTobacco_BeforeCollege <- rowSums(substance[c(
"TotalDaysUsedTobacco_6thTo8thGrade",
"TotalDaysUsedTobacco_9thGrade",
"TotalDaysUsedTobacco_10thGrade",
"TotalDaysUsedTobacco_11thGrade",
"TotalDaysUsedTobacco_12thGrade",
"TotalDaysUsedTobacco_SummerAfterHighSchool"
)], na.rm = T)

substance$DaysBinge_College <- rowSums(substance[c(
"DaysBingeDrinking_Female_FreshmanYear",
"DaysBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysBingeDrinking_Female_ThisAcademicYear",
"V1_DaysBingeDrinking_Female_Past30Days",
"DaysBingeDrinking_Male_FreshmanYear",
"DaysBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysBingeDrinking_Male_ThisAcademicYear",
"V1_DaysBingeDrinking_Male_Past30Days"
)], na.rm = T)

substance$DaysExtremeBinge_College <- rowSums(substance[c(
"DaysExtremeBingeDrinking_Female_FreshmanYear",
"DaysExtremeBingeDrinking_Female_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Female_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Female_Past30Days",
"DaysExtremeBingeDrinking_Male_FreshmanYear",
"DaysExtremeBingeDrinking_Male_SummerAfterFreshmanYear",
"DaysExtremeBingeDrinking_Male_ThisAcademicYear",
"V1_DaysExtremeBingeDrinking_Male_Past30Days"
)], na.rm = T)

substance$DaysUsedCannabis_College <- rowSums(substance[c(
"TotalDaysUsedCannabis_FreshmanYear",
"TotalDaysUsedCannabis_SummerAfterFreshmanYear",
"TotalDaysUsedCannabis_ThisAcademicYear",
"V1_TotalDaysUsedCannabis_Past30Days"
)], na.rm = T)

substance$DaysUsedTobacco_College <- rowSums(substance[c(
"TotalDaysUsedTobacco_FreshmanYear",
"TotalDaysUsedTobacco_SummerAfterFreshmanYear",
"TotalDaysUsedTobacco_ThisAcademicYear",
"V1_TotalDaysUsedTobacco_Past30Days"
)], na.rm = T)

substance$DaysBinge <- substance$DaysBinge_BeforeCollege + substance$DaysBinge_College
substance$DaysExtremeBinge <- substance$DaysExtremeBinge_BeforeCollege + substance$DaysExtremeBinge_College
substance$DaysUsedCannabis <- substance$DaysUsedCannabis_BeforeCollege + substance$DaysUsedCannabis_College
substance$DaysUsedTobacco <- substance$DaysUsedTobacco_BeforeCollege + substance$DaysUsedTobacco_College
names(substance)[1] <- "id"

lifetimesubstance <- substance[c(1, 90:93)]
saveRDS(object = lifetimesubstance, file = "~/Documents/oleary/rest/lifetimesubstance.rds")
```

### Personality Variables

```{r}
personality_scales <- c(
"Subject_ID",
"V1_BIS_Motor_Combined",
"V1_BIS_Attentional_Combined",
"V1_BIS_Nonplanning_Combined",
"V1_SSS_Boredom_Susceptibility",
"V1_SSS_Thrill_Adventure_Seeking",
"V1_SSS_Disinhibition_without_substance_questions",
"V1_SSS_Experience_Seeking_without_substance_questions",
"V1_SSS_Overall_Score_no_substances")

personality <- spss[personality_scales][,1:8]
names(personality)[1] <- "id"

saveRDS(object = personality, file = "~/Documents/oleary/rest/personality.rds")
```


```{r}
behav <- merge(lifetimesubstance, yaacq, all = T)
#behav <- merge(merge(lifetimesubstance, yaacq, all = T), personality, all = T)
saveRDS(object = behav, file = "~/Documents/oleary/rest/behav.rds")
```


```{r}
rs_connectivity <- readRDS("~/Documents/oleary/rest/91rsFC.rds")
behav <- readRDS("~/Documents/oleary/rest/behav.rds")
nuisance_covs <- readRDS("~/Documents/oleary/rest/covariates.rds")
```

Print number of rows of each loaded dataframe

```{r}
c('rs_connectivity' = nrow(rs_connectivity), 
  'behav' = nrow(behav), 
  'nuisance_covs' = nrow(nuisance_covs)) 
```

Are subjects in data frames in the same order?

```{r}
all(rs_connectivity$id == nuisance_covs$id)
all(rs_connectivity$id == behav$id) 
```

Throw away the id column, because it is not needed anymore

```{r}
rs_connectivity <- rs_connectivity[, names(rs_connectivity) != 'id']
behav <- behav[, names(behav) != 'id']
nuisance_covs <- nuisance_covs[, names(nuisance_covs) != 'id'] 
```


## Summary statistics

Summary of all subjects

```{r}
summary(nuisance_covs) 
```


## Regress out age and sex 

```{r}
residual_rs_connectivity <- matrix(NA,
                                   nrow = nrow(rs_connectivity), 
                                   ncol = ncol(rs_connectivity))

for (i in 1:ncol(rs_connectivity)) {
  fit <- lm(rs_connectivity[,i] ~ age + factor(sex),
            data = nuisance_covs, na.action = na.exclude)
  residual_rs_connectivity[,i] <- residuals(fit)
}

residual_rs_connectivity <- data.frame(residual_rs_connectivity)
names(residual_rs_connectivity) <- names(rs_connectivity)

```


# 1. Canonical correlation analysis

## Feature selection and CCA function

Here we create a function that first selects resting state features (X) with the highest spearman correlation with any of behav symptoms (Y) and then fits and returns a CCA model. This function will be used to compute canonical correalotions and also later for permutation test and cross-validation.

```{r}
select_and_cca_fit <- function(X, Y, n_selected_vars){
  library(candisc)
  #select
  correlations <- cor(Y, X, method = "spearman")
  correlations <- apply(correlations, 2, function(x){max(abs(x))})
  corr.threshold <- sort(correlations, decreasing = T)[n_selected_vars]
  selected.X <- correlations >= corr.threshold
  selected.X <- X[,selected.X]
  #cca fit
  cca_model <- candisc::cancor(selected.X, Y)
  #return fitted model containing canonical correlations and wilks lambdas
  return(cca_model)
}

```

## Canonical correlations

Fit the feature selection and CCA model, selecting 150 features and print all canonical correlations

```{r}
n_selected_vars <- 91
cca_model <- select_and_cca_fit(residual_rs_connectivity, 
                                behav, 
                                n_selected_vars)
cca_model$cancor
Wilks(cca_model)
```

Create a function to compute canonical variates

```{r}
predict.cancor <- function(cancor.obj, X, Y){
  X_pred <- as.matrix(X) %*% cancor.obj$coef$X
  Y_pred <- as.matrix(Y) %*% cancor.obj$coef$Y
  XY_pred <- list(X_pred, Y_pred)
  names(XY_pred) <- c("X_pred", "Y_pred")
  return(XY_pred)
}
```

Visualize canonical correlations 

```{r}
canonical.variates <- predict.cancor(cca_model, 
                               residual_rs_connectivity[,cca_model$names$X], 
                               behav)
cca_y_loadings <- cor(behav, canonical.variates$Y_pred)

par(mfrow=c(1,2))
plot(canonical.variates$X_pred[,1], 
     canonical.variates$Y_pred[,1],
     bty='n',
     xlab='Connectivity canonical variate 1',
     ylab='Behavioral canonical variate 1')
text(-1, 1,  bquote(r^2 == .(round(cca_model$cancor[1], 2))))

plot(canonical.variates$X_pred[,2], 
     canonical.variates$Y_pred[,2],
     bty='n',
     xlab='Connectivity canonical variate 2',
     ylab='Behavioral canonical variate 2')
text(-2, 2,  bquote(r^2 == .(round(cca_model$cancor[2], 2))))

```

## Permutation test

First get test statistics (canonical correlations and Wilks lambdas) from the real model 

```{r}
real_model <- cca_model
real_results_cancor <- real_model$cancor
real_results_wilks <- Wilks(real_model)$"LR test stat"
```

Obtain null distribution of test statistics by permuting rows of behav data

```{r}
library(permute)
library(doMC)
registerDoMC(cores=4) # to run it multicore

nperms = 1999
set.seed(123)

# shuffle 
shuffled_indexes <- sapply(1:nperms, function(x){
                           shuffle(1:nrow(behav))})

null_results <- foreach(i=1:nperms) %dopar% {
  null_model <- select_and_cca_fit(residual_rs_connectivity,
                                    behav[shuffled_indexes[,i],], 
                                    n_selected_vars)
  #return canonical correlations and wilks lambdas
  list(null_model$cancor, Wilks(null_model)$"LR test stat")
}

# transform null results lists to data frame
null_dist_cancor <- lapply(null_results, function(x){return(x[[1]])})
null_dist_wilks <- lapply(null_results, function(x){return(x[[2]])})
null_dist_cancor <- as.data.frame(do.call(rbind, null_dist_cancor))
null_dist_wilks <- as.data.frame(do.call(rbind, null_dist_wilks))

get_pval <- function(real, null_dist, better="smaller"){
  if (better == "smaller"){
    rank <- sum(real < null_dist) + 1
  }
  if (better == "bigger"){
    rank <- sum(real > null_dist) + 1
  }
  pval <- rank / (length(null_dist) + 1)
  return(pval)
}

pvals_cancor <- mapply(function(real, null_dist){
                         get_pval(real, null_dist, better="smaller")},
                       real_results_cancor,
                       null_dist_cancor)
pvals_wilks <- mapply(function(real, null_dist){
                        get_pval(real, null_dist, better="bigger")},
                       real_results_wilks,
                       null_dist_wilks)
```

Print p-values

```{r}
print(cbind("component"=1:length(pvals_cancor), pvals_cancor, pvals_wilks))
```

Visualize null distributions and p-values for first two canonical correlations

```{r}
par(mfrow=c(2,1))
for (i in 1:2){
  hist(null_dist_cancor[,i], breaks = 25, main = paste("Null dist corr CV", i),
       xlim=c(0.7,1))
  abline(v=real_results_cancor[i], col="red")
}
```


## Stability of canonical loadings

Create function that performes leave-one-out jackknife procedure to get uncertainity of canonical loadings taking into an account uncertainity caused by feature selection. 

Jackknife repeatedly leaves one subject out and then performs the feature selection and CCA procedure in the same way as above. 

```{r}
njack <- nrow(residual_rs_connectivity)
jack_res <- foreach(i=1:njack) %dopar% {
  model <- select_and_cca_fit(residual_rs_connectivity[-i,],
                               behav[-i,],
                               n_selected_vars)
  selected.vars <- model$names$X
  prediction <- predict.cancor(model, 
                               residual_rs_connectivity[i, selected.vars], 
                               behav[i,])
  list(prediction, model)
}
```

run jackknife

```{r}
jack.results <- lapply(jack_res, function(x){return(x[[1]])})
jack.X <- lapply(jack.results, function(x){return(x[[1]])})
jack.X <- as.data.frame(do.call(rbind, jack.X))
jack.Y <- lapply(jack.results, function(x){return(x[[2]])})
jack.Y <- as.data.frame(do.call(rbind, jack.Y))
```

get loadings from saved jackknife models

```{r}
jack_models <- lapply(jack_res, function(x){return(x[[2]])})
jack.loadings <- lapply(jack_models, function(model){
  return(model$structure$Y.yscores[,1])})
jack.loadings <- as.data.frame(do.call(rbind, jack.loadings))
```

plot distribution of canonical loadings across all jackknife models

```{r, message=F}
library(reshape2)
melted.loadings <- melt(jack.loadings)

par(mfrow=c(1,2), las=1, mai=c(1.02, 1.3, 0.82, 0.42))
boxplot(abs(value) ~ variable, data=melted.loadings, horizontal=T)
plot(abs(melted.loadings$value), 
     jitter(as.numeric(melted.loadings$variable)), 
     pch='.')
```

# 3. Software enviroment
```{r}
sessionInfo()
```
